\chapter{Advice}
\label{cha:advice}

\section{General Advice}
\label{sec:general-advice}

\begin{itemize}
\item Do about four courses per term.
\item Start around six courses per term, narrow it down.
\item Understanding lecture $N$ before lecture $N+1$.
\item Enjoyment of different courses.
\end{itemize}

\section{Examples Classes}
\label{sec:examples-classes}

\begin{itemize}
\item Gives a sheet of problems - in two weeks, we will go through the questions.
\item Very important to attempt the questions.
\end{itemize}

\section{Student Groups}
\label{sec:student-groups}

\begin{itemize}
\item Group of students getting together to support each other.
\end{itemize}

\section{Advisor}
\label{sec:advisor}

\begin{itemize}
\item October 16 - Careers, USA PhD's, funding.
\item October 30 PhD's at Cambridge and in the UK.
\item Open Days - Statistics Lab, Cambridge Centre of Analysis
  (\textsc{cca}).
\end{itemize}

\section{Exams}
\label{sec:exams}

\begin{itemize}
\item 17-19 units taken
\item 17/18/19 cannot hurt to take.
\item Start of Easter term is where the decision is made.
\end{itemize}

\section{Essays}
\label{sec:essays}

\begin{itemize}
\item Read a bunch of mathematical papers, present the papers at a
  excellent level.
\item \textsc{e.g.} paper gives definition, essay gives examples.
\item People tend to enjoy the essays.
\end{itemize}

\section{Student Support}
\label{sec:student-support}

\begin{itemize}
\item Tuesday 15 October - Probability and Measure catch-up lectures.
\end{itemize}

Part III Cafe
\begin{itemize}
\item Mon/Wed/Fri 4/5/4
\end{itemize}

Study Groups

\begin{itemize}
\item Discuss lectures with you colleagues
\item examples/questions
\item Organize yourselves
\end{itemize}

Wednesday Talks

\begin{itemize}
\item Every Wednesday 4:45P.
\item Talks by different people of topics
\item Helps for applying to PhD's
\item How to propose a talk
\item How to read a paper
\end{itemize}

Calendars

\begin{itemize}
\item \url{cugms.soc.sucf.net} - separate page for Part III students
\item \url{www.maths.cam.ac.uk/facultyboard/calendars} - Google
  Calendar for Part III mathematics
\end{itemize}


%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "master"
%%% End: 
