\chapter{Continuous Time Models}
\label{cha:cont-time-models}

In discrete time, we had $X_{t} - X_{t-1} = H_{t} \cdot (P_{t} -
P_{t-1}) - c_{t}$.  For continuous time, we replace this with $dX_{t}
= H_{t} dP_{t} - c_{t} dt$

A state price density is some stochastic process
$Y$ with $Y_{t} > 0$ and $YP$ is a martingale

\begin{lem}
  If $t \mapsto X_{t}(\omega)$ is differentiable and $X$ is a
  martingale then $X$ is constant.
\end{lem}

This can make a pricing theory quite boring!

\newcommand{\ito}{It\^o\xspace}
\section{Diversion into Stochastic Calculus}
\label{sec:divers-into-stoch}

\begin{defn}
  \label{defn:continuous_time:1}
  A (standard scalar) Brownian motion is a process $W = (W_{t})_{t
    \geq 0}$ such that
  \begin{enumerate}
  \item $W_{0}(\omega) = 0$ for all $\omega$.
  \item $t \mapsto W_{t}(w)$ is continuous for all $\omega$
  \item For any $0 \leq t_{0} < t_{1} < \dots < t_{n}$, the increments
    $W_{t_{1}} - W_{t_{0}}, \dots, W_{t_{n}} - W_{t_{n-1}}$ are
    independent, with $W_{t} - W_{s} \sim N(0, |t-s|)$.
  \end{enumerate}
\end{defn}

\begin{thm}
  \label{defn:continuous_time:1}
  The Brownian motion exists (Weiner, 1923).
\end{thm}

Consider a filtration $(\mathcal{F}_{t})$ with the property that
$W_{t} - W_{s}$ is independent of $\mathcal{F}_{s}$, $0 \leq s \leq
t$. Our technical assumptions are usual conditions - $\mathcal{F}_{t}
= \cap_{\epsilon > 0} \mathcal{F}_{t + \epsilon}$ (right-continuity),
$\mathcal{F}_{0}$ contains all $\Prob$-null sets.

\begin{defn}
  \label{defn:continuous_time:2}
  A \textbf{simple predictable process} is of the form
  \begin{equation}
    \label{eq:31}
    \alpha_{t}(\omega) = \sum_{i=1}^{n} \I{(t_{i-1}, t_{i})}
    a_{i}(\omega),
  \end{equation}
  where $0 \leq t_{0} < \dots < t_{n}$, each $a_{i}$ is a bounded
  $\mathcal{F}_{t_{i-1}}$-measurable random variable.
\end{defn}

\begin{remark}
  $\alpha$ is left-continuous, piecewise-constant, and adapted.
\end{remark}

\begin{defn}
  \label{defn:continuous_time:3}
  \begin{equation}
    \label{eq:30}
    \int_{0}^{\infty} \alpha_{s} dW_{s} = \sum_{i=1}^{n} a_{i}
    (W_{t_{i}} - W_{t_{i-1}})
  \end{equation} where $\alpha$ is a simple predictable process.
\end{defn}

\begin{defn}
  \label{defn:continuous_time:4}
  The predictable $\sigma$-algebra on $[0, \infty) \times \Omega$ is
  generated by $(s, t] \times A$ where $A \in \mathcal{F}_{s}$.

  This is the smallest $\sigma$-algebra for which simple predictable
  processes are measurable.

  A process measurable with respect to the predictable
  $\sigma$-algebra is called \textbf{predictable}.
\end{defn}

\begin{remark}
  If $\alpha$ is left-continuous and adapted, it is predictable.
\end{remark}

\begin{proposition}[\ito's isometry]
  If $\alpha$ is simple and predictable, then
  \begin{equation}
    \label{eq:32}
    \E{\left(\int_{0}^{\infty}\alpha_{s} dW_{s} \right)^{2}} =
    \E{\int_{0}^{\infty} \alpha_{s}^{2} ds}
  \end{equation}


  Thus, the isometry $I$ from simple predictable process to square
  integrable random variables on $L^{2}(\Omega, \mathcal{F}, \Prob)$
  (which is complete) defined by
  \begin{equation}
    \label{eq:33}
    I(\alpha) = \int_{0}^{\infty} \alpha_{s} dW_{s}
  \end{equation}
\end{proposition}

\begin{proof}
  \begin{align}
    \label{eq:34}
    \left( \int \alpha dW \right)^{2} &= \left(\sum a_{i} \Delta
      W_{i}\right)^{2} \\
    &= 2 \sum_{j < i} a_{j} a_{i} \Delta W_{j} \Delta W_{i} + \sum
    a_{i}^{2} (\Delta W_{i})^{2}
  \end{align}

  Note that $\E{\sum a_{i}^{2 (\Delta W_{i})^{2}}} = ...$

  \todo{Finish this proof}
\end{proof}

\begin{defn}
  \label{defn:continuous_time:5}
  Suppose $\E{\int_{0}^{\infty} (\alpha^{k}_{s} - \alpha_{s})^{2} ds}
  \rightarrow 0$, where each $\alpha^{k}$ is simple and predictable.
  Then
  \begin{equation}
    \label{eq:35}
    \int_{0}^{\infty} \alpha_{s} dW_{s} = \lim_{L^{2}}
    \int_{0}^{\infty} a_{s}^{k} dW_{s}
  \end{equation}
\end{defn}

\begin{thm}
  \label{defn:continuous_time:6}
  If $\alpha$ is predictable and $\E{\int_{0}^{t} \alpha_{s}^{2} ds} <
  \infty$ for all $t$, there exists a continuous martingale $X$ such
  that $X_{t} = \int_{0}^{\infty} \alpha_{s} \I{s \leq t} dW_{s}$.

\end{thm}

For notation, we represent $X_{t}$ as $\int_{0}^{t} \alpha_{s}
dW_{s}$.  Note that $\E{X_{t}} = 0$ and $\E{X_{t}^{2}} = \int_{0}^{t}
\alpha_{s}^{2} ds$.

\begin{defn}[Localization]
  \label{defn:continuous_time:7}
  Suppose $\alpha$ is predictable and $\int_{0}^{t} \alpha_{s}^{2} ds
  < \infty$ almost surely for all $t$.  Let $\tau_{n} = \inf \{ t \geq
  0 | \int_{0}^{t} \alpha_{s} ds > n \}$.

  Let $\alpha^{(n)}_{t} = \alpha_{t} \I{t \leq \tau_{n}}$, so
  $\int_{0}^{t} \alpha_{s}^{(n)} dW_{s}$ is well-defined by the
  $L^{2}$ theory, since $\E{\int_{0}^{t} (\alpha_{s}^{(n)})^{2} ds}
  \leq N \leq \infty$ as $\int_{0}^{t} \alpha_{s}^{2} ds < \infty$
  almost surely as $\tau_{n} \uparrow \infty$.
\end{defn}

\begin{notation}
  $\int_{0}^{t} \alpha_{s} dW_{s}$ as $\int_{0}^{t} \alpha_{s}^{(N)}
  dW_{s}$ on $\{ t \leq \tau_{n} \}$.
\end{notation}

\begin{thm}
  \label{defn:continuous_time:8}
  If $\alpha$ is adapted and continuous, then $\int_{0}^{t} \alpha_{s}
  dW_{s}$ is defined for all $t \geq 0$ - since $t \mapsto
  \alpha_{t}(\omega)$ is continuous, $\alpha$ is bounded on $[0, t]$
  for each $\omega$, and so $\int_{0}^{t} \alpha_{s} ds < \infty$
  almost surely.

  If $X_{t} = \int_{0}^{t} \alpha_{s} dW_{s}$, then $X$ is a
  continuous local martingale, since $X^{(n)} = (X_{t \wedge
    \tau_{n}})_{t} \geq 0$ is a true martingale, where $\tau_{n} = \inf
  \{ \tau \geq 0, \int_{0}^{t} \alpha_{s} ds \geq N \}$.
\end{thm}



\section{\ito's Formula}
\label{sec:itos-formula}

\begin{defn}
  \label{defn:continuous_time:9}
  An \ito process $X$ is of the form
  \begin{equation}
    \label{eq:36}
    X_{t} = X_{0} + \int_{0}^{t} \alpha_{s} dW_{s} + \int_{0}^{t}
    \beta_{s} ds
  \end{equation}
  such that $\alpha, \beta$ are predictable and $\int_{0}^{t}
  \alpha_{s} ds < \infty$ and $\int_{0}^{t} |\beta_{s}| ds < \infty$
  for all $t$.
\end{defn}

\begin{thm}
  \label{defn:continuous_time:10}
  If $X$ is an \ito process and $f \in C^{2}$, then $f(X)$ is an \ito
  process.  In fact,
  \begin{equation}
    \label{eq:37}
    f(X_{t}) = f(X_{0}) + \int_{0}^{t} f'(X_{s}) \alpha_{s} dW_{s} +
    \int_{0}^{t} \left( f'(X_{s}) \beta_{s} + \underbrace{\frac{1}{2}f''(X_{s})
      \alpha_{s}^{2}}_{\text{\ito's correction}} \right) ds
  \end{equation}
\end{thm}

\begin{exmp}
  \label{defn:continuous_time:11}
  $f(x) = x^{2}$. Then
  \begin{align}
    \label{eq:38}
    W_{t}^{2} = \int_{0}^{t} 2 W_{s} dW_{s} + t \\
    \E{W_{t}^{2}} = \E{\int_{0}^{t} 2 W_{s} dW_{s}} + t
  \end{align} and the first term is zero as it is a martingale.

  This follows from
  \begin{equation}
    \label{eq:39}
    \E{\int_{0}^{t} W_{s}^{2} ds} = \int_{0}^{t} s ds =
    \frac{t^{2}}{2} < \infty
  \end{equation} so $\int_{0}^{t} W_{s} dW_{s}$ is a martingale.
\end{exmp}

\begin{thm}
  \label{defn:continuous_time:12}
  Let $X$ be an \ito process.  Fix $t > 0$.  Then
  \begin{equation}
    \label{eq:40}
    \lim_{n \rightarrow \infty} \sum_{k=1}^{n} \left(X_{\frac{tk}{n}}
      - X_{\frac{t(k-1)}{n}}\right)^{2} = \int_{0}^{t} \alpha_{s}^{2} ds
  \end{equation}
\end{thm}

\begin{notation}
  \begin{equation}
    \label{eq:41}
    \IP{ X }_{t} = \int_{0}^{t} \alpha_{s} ds
  \end{equation} is called the quadratic variation of $X$.
\end{notation}

\begin{thm}[\ito's formula]
  \label{defn:continuous_time:13}
  In integral form,
  \begin{equation}
    \label{eq:42}
    f(X_{t}) = f(X_{0}) + \int_{0}^{t} f'(X_{s}) dX_{s} + \frac{1}{2}
    \int_{0}^{t} f''(X_{s}) d \IP{ X }_{s}
  \end{equation}

  In differential form,
  \begin{equation}
    \label{eq:43}
    df(X_{t}) = f'(X_{t}) dX_{t} + \frac{1}{2} f''(X_{t}) d\IP{ X }_{t}
  \end{equation}
\end{thm}

Morally, the idea is to take Taylor expansion around $f(X_{t})$.

\begin{thm}[\ito's formula, multidimensional version]
  \label{defn:continuous_time:14}
  let $X, Y$ be \ito processes. Then the quadratic covariation
  \begin{align}
    \label{eq:44}
    \IP{ X, Y }_{t} &= \lim_{n \rightarrow \infty}
    \sum_{k=1}^{n} (X_{\frac{tk}{n}} -
    X_{\frac{t(k-1)}{n}})(Y_{\frac{tk}{n}} -
    Y_{\frac{t(k-1)}{n}}) \\
    &= \frac{1}{2} \IP{ X + Y }_{t} - \IP{ X }_{t} -
    \IP{ Y }_{t}
  \end{align}
\end{thm}

\begin{proposition}
  The quadratic covariance satisfies the following properties:
  \begin{enumerate}
  \item (Bilinear, symmetric)
    \begin{equation}
      \label{eq:45}
      \IP{ aX + bY, Z } = a \IP{ X, Z } + b \IP {Y, Z} = \IP{ Z, aX + bY }
    \end{equation}
  \item If $X_{t} = X_{0} + \int_{0}^{t} \beta_{s} ds$ then $\IP{X,
      Y}_{t} = 0$ for any \ito process $Y$.
  \item Let $W^{1}, W^{2}$ be two independent Brownian motions.  Then
    $\IP{ W^{1}, W^{2} }_{t} = 0$.
  \item
    \begin{equation}
      \label{eq:46}
      \IP{\int_{0}^{t} \alpha_{s} dW_{s}, \int_{0}^{t} \beta_{s}
        dW_{s}} = 
      \int_{0}^{t} \alpha_{s} \beta_{s} ds
    \end{equation}
  \end{enumerate}
\end{proposition}

Let $X$ be an $n$-dimensional \ito process, and $f \in C^{2}(\R^{n}
\rightarrow \R)$.  Then
\begin{equation}
  \label{eq:47}
\end{equation}

\todo{Fill in this multivariate \ito's result}

In finance there are state price densities $\Rightarrow$ equivalent
martingale measures.  How to do computations under equivalent changes
of measure?

Let $W$ be an $n$-dimensional BM with $W = (W^{1}, \dots, W^{m})$
where $W^{i}$ are independent standard Brownian motions.  Let $\alpha$
be an $n$-dimensional predictable process and $\int_{0}^{t} \| a_{s}
\|^{2} ds < \infty$, and let

\begin{equation}
  \label{eq:48}
  Z_{t} = e^{\int_{0}^{t} \alpha_{s}
    dW_{s} - \frac{1}{2} \int_{0}^{t} \| \alpha_{s} \|^{2} ds}.
\end{equation}

\begin{proposition}
  $Z$ satisfies the following properties:
  \begin{enumerate}
  \item $Z$ is a local martinagle.
  \item $Z$ is a supermartingale.
  \item If $\E{Z_{T}} = 1$ for some $T > 0$ (non-random), then
    $(Z_{t})_{0 \leq t \leq T}$ is a true martingale.
  \end{enumerate}
\end{proposition}

\begin{proof}
  Let $dX_{t} = \alpha_{t} \cdot dW_{t} - \frac{1}{2} \| \alpha_{t}
  \|^{2} dt$, $X_{0} = 0$. Let $f(x) = e^{x}$.  Then
  \begin{align}
    \label{eq:49}
    dZ_{t} = df(X_{t}) = f'(X_{t}) dX_{t} + \frac{1}{2} f''(X_{t})
    d\IP{ X }_{t}
  \end{align}
  Note that
  \begin{align}
    \label{eq:50}
    d \IP{ X }_{t} &= d \IP{\sum_{i=1}^{m} \int_{0}^{t} \alpha_{s}^{2} dW_{s}^{2}}_{t} \\
    &= d \sum_{i, j} \IP{\int \alpha_{s}^{i} dW^{i}_{s}, \int \alpha_{s}^{j} dW^{j}}_{t} \\
    &= \sum (\alpha_{t}^{i})^{2} dt \\
    &= \| \alpha_{t} \|^{2} dt
  \end{align}

  Then
  \begin{align}
    \label{eq:51}
    dZ_{t} = Z_{t} \left(\alpha_{t} \cdot dW_{t} - \frac{1}{2} \|
      \alpha_{t} \|^{2} dt \right) + \frac{1}{2} Z_{t} \| \alpha_{t}
    \|^{2} dt = Z_{t} \alpha_{t} dW_{t}.
  \end{align}

  Thus
  \begin{equation}
    \label{eq:52}
    Z_{t} = 1 + \int_{0}^{t} Z_{s} \alpha_{s} \cdot dW_{s}
  \end{equation} and so $Z$ is a stochastic integral, and hence a
  local martingale.

  $Z_{t} > 0$ almost surely, so non-negative local martingales are
  supermartingales by Fatou's lemma.

  $Z$ is a supermartinagle and $\E{Z_{T}} = Z_{0}$, and so $(Z_{t})_{0
  \leq t \leq T}$ is a martingale (pigeonhole principle).
\end{proof}

\begin{thm}[Cameron-Martin-Girsanov theorem]
  \label{defn:continuous_time:15}
  Let $Z$be as before and assume $\E{Z_{T}} = 1$for some $T > 0$.
  Define an equivalent martingale measure $\Q$ by Radon-Nikodym
  density
  \begin{equation}
    \label{eq:53}
    \frac{d\Q}{d\Prob} = Z_{t}
  \end{equation}  Let $\hat W_{t} = W_{t} - \int_{0}^{t} \alpha_{s}
  ds$.  Then $\hat W$ is a $\Q$-Brownian motion.
\end{thm}

\begin{thm}[Martingale representation theorem]
  \label{defn:continuous_time:16}
  Let $W$ be an $m$-dimensional Brownian motion generating the
  filtration $(\mathcal{F}_{t})_{t \geq 0}$.
  Let $X$ be a continuous local martingale.  Then there exists a
  predictiable $\alpha$ with $\int_{0}^{t} \| \alpha_{s} \|^{2} ds <
  \infty$ almost surely for all $t$ such that $X_{t} = X_{0} +
  \int_{0}^{t} \alpha_{s} dW_{s}$.

  If $X_{t} > 0$ a.s. for all $t$, then there exists a predictable
  process $\beta$ with $\int_{0}^{t} \| \beta_{s} \|^{2} ds < \infty$
  for all $t$ such that
  \begin{equation}
    \label{eq:54}
    X_{t} = X_{0} e^{\int_{0}^{t} \beta_{s} dW_{s} - \frac{1}{2}
      \int_{0}^{t} \| \beta_{s} \|^{2} ds}
  \end{equation}
\end{thm}

\begin{thm}[Levy's characterization theorem]
  \label{defn:continuous_time:17}
  Let $X$ be a continuous local martingale (in any filtration
  satisfying the usual conditions) such that its quadratic variation
  $\IP{ X }_{t} = t$.  Then $X$ is a Brownian motion.
\end{thm}


\section{Arbitrage Theory in Continuous Time}
\label{sec:arbitr-theory-cont}

Recall that in discrete time,
\begin{align}
  \label{eq:55}
  X_{t} = H_{t} \cdot P_{t} = H_{t+1} \cdot P_{t} - c_{t+1} \\
  X_{t+1} = H_{t+1} \cdot P_{t+1} \Rightarrow X_{t+1} - X_{t} =
  H_{t+1} \cdot (P_{t+1} - P_{t}) - c_{t+1}
\end{align}

The setup is as follows:
\begin{enumerate}
\item $P$ is an $m$-dimensional \ito process.
\end{enumerate}

\begin{defn}
  \label{defn:continuous_time:18}
  A self-financing investment/consumption strategy $(H, c)$ is a pair
  of predictable processes such that $c_{t} \geq 0$ for all $t$ ,
  $\int_{0}^{t} \sum (H_{s}^{i})^{2} d \IP{ P^{i} }>_{s} <
  \infty$ for all $t$, and
  \begin{equation}
    \label{eq:56}
    H_{t} \cdot P_{t} = H_{0} \cdot P_{0} + \int_{0}^{t} H_{s} \cdot
    dP_{s} - \int_{0}^{t} c_{s} ds
  \end{equation}
\end{defn}

\begin{defn}[Incomplete]
  \label{defn:continuous_time:19}
  An arbitrage is an investment/consumption strategy $(H, c)$ such
  that $X_{0} = X_{T} = 0$ and $\Prob{\int_{0}^{T} c_{s} ds > 0} > 0$
  for some non-random $T > 0$
\end{defn}

This definition is flawed.

\begin{exmp}[Doubling strategies]
  \label{defn:continuous_time:20}
  Consider the discrete-time model $P = (1, S_{t})$ where $S_{t} =
  \xi_{1} + \dots + \xi_{t}$ where $\xi_{i}$ are IID with
  $\Prob{\xi_{i} = \pm 1} = \frac{1}{2}$.
\end{exmp}

Consider a price vector $P = (1, W)$ with $W$ a Brownian motion . Let
$X_{t} = \int_{0}^{t} \pi_{s} dW_{s}$, and let $f: [0, 1] \rightarrow
[0, \infty]$ an increasing bijection with inverse $f^{-1}$.  For
example, $f(t) = \frac{t}{1-t}$ with $f^{-1}(u) = \frac{u}{1+u}$.
Consider
\begin{align}
  \label{eq:57}
  Z_{u} = \int_{0}^{f^{-1}(u)} \sqrt{f'(s)} dW_{s}
\end{align}

Then
\begin{align}
  \label{eq:58}
  \IP{ Z }_{u} = \int_{0}^{f^{-1}(u)} f'(s) ds = u
\end{align} which implies $Z$ is a Brownian motion by Levy's
characterization.  Let $\tau = \in \{ u \geq 0 : Z_{u} > K \}$ where
$K > 0$ is a constant.  Let $\pi_{t} = \sqrt{f'(t)} \I{t \leq
  f^{-1}(\tau)}$.  Note that $\int_{0}^{1} \pi^{2}_{s} ds =
\int_{0}^{f^{-1}(\tau)} f'(s) ds = \tau < \infty$.  So $\int_{0}^{t}
\pi_{s} dW_{s}$ makes sense for all $t \leq 1$.  Let $X_{t} = \pi_{s}
dW_{s}$, with $X_{1} = \int_{0}^{f^{-1}(\tau)} \sqrt{f'(s)} dW_{s} =
Z_{\tau} = K > 0$.  $X$ is a local martingale since it is a stochastic
integral, but $\E{X_{1}} - K \neq X_{0} = 0$.

\begin{defn}
  \label{defn:continuous_time:21}
  An investment/consumption strategy $(H, c)$ is $L$-admissible if
  $X_{t}(H, c) \geq -L_{t}$ for all $t$ a.s. where $L$ is given
  non-negative adapted process.

  For most cases, $L = 0$.
\end{defn}

\begin{defn}
  \label{defn:continuous_time:22}
  A state price density is a positive \ito process such that $(Y_{t}
  P_{t})_{t \geq 0}$ is a local martingale.
\end{defn}

\begin{thm}
  \label{defn:continuous_time:23}
  If there exists a state price density such that $YL$ is uniformly
  integrable, then there is no arbitrage among $L$-admissible
  self-financing investment/consumption strategies.
\end{thm}

\begin{remark}
  Recall that $(Z_{t})_{t \geq 0}$ is uniformly integrable if and only
  if
  \begin{equation}
    \label{eq:59}
    \lim_{k \rightarrow \infty} \sup_{t \geq 0} \E{|Z_{t}|\I{Z_{t \geq
        k}}} = 0
  \end{equation}
\end{remark}

\begin{remark}
  If $(Z_{t})_{0 \leq t \leq T}$ is a martingale then $(Z_{t})_{0 \leq t
  \leq T}$ is uniformly integrable ($T < \infty$ not random.)
\end{remark}

\begin{remark}
  If $\sup_{t \geq 0} \E{|Z_{t}|^{p}} < \infty$ for some $p > 1$ then
  $(Z_{t})_{t \geq 0}$ is uniformly integrable.
\end{remark}

\begin{remark}
  If $Z_{n} \rightarrow Z_{\infty}$ a.s. and $(Z_{n})_{n \geq 1}$ is
  UI then $\E{|Z_{n} - Z_{\infty}|} \rightarrow 0$.
\end{remark}

\begin{proposition}
  Let $(H, c)$ be a self financing stragey and $X_{t} = H_{t} \cdot
  P_{t}$ so that $dX_{t} = H_{t} \cdot dP_{t} - c_{t} dt$.  Let $Y$ be
  an \ito process.  Let $Y$ be an \ito process.  Then
  \begin{equation}
    \label{eq:60}
    d(X_{t} Y_{t} ) = H_{t} \cdot (dY_{t}P_{t}) - Y_{t} c_{t} dt.
  \end{equation}
\end{proposition}

\begin{proof}
  Since $dX = H \cdot dP - c dt$, then
  \begin{align}
    \label{eq:61}
    d \IP{ X, Y } = \sum_{i=1}^{n} h^{i} d \IP{P^{i}, Y^{i}}
  \end{align}

  By \ito's formula,
  \begin{align}
    \label{eq:62}
    d(XY) &= XdY + YdX + d<X, Y>  \\
    &= H \cdot P dY + Y(H \cdot dP - c dt) + \sum H^{i} d \IP{P^{i},
      Y^{i}} \\
    &= \sum H^{i}(P^{i} dY + Y dP^{i} + d \IP{ P^{i}, Y }) -
    Y c dt \\
    &= \sum H^{i} d(P^{i} Y) - Yc dt
  \end{align}
\end{proof}

\begin{defn}
  \label{defn:continuous_time:24}
  A continuous, adapted process $(Z_{t})_{t \geq 0}$ is of class
  $\mathcal{D}$ (Doob) if $\{ Z_{\tau} | \text{$\tau$ stopping times}
  \}$ is uniformly integrable.
\end{defn}

\begin{remark}
  If $\E{\sup_{t \geq 0} |Z_{t}|} < \infty$, then $(Z_{t})_{t \geq 0}$
  is of class $\mathcal{D}$.
\end{remark}

\begin{thm}
  \label{defn:continuous_time:25}
  If $YL$ is of class $D$ (at least locally), then there is no arbitrage.
\end{thm}

\begin{thm}
  \label{defn:continuous_time:26}
  If there exists a state price density $Y$ such that $YL$ is of class
  $D$ locally, then there are no $L$-admissible arbitrages.

  Class $D$ locally means $\{ \text{$Z_{\tau \wedge t}$ - $\tau$ a
    stopping time is $UI \forall t \geq 0$} \}$.
\end{thm}

\begin{proof}
  \begin{align}
    \label{eq:63}
    \int_{0}^{t} H_{s} \cdot d(X_{s} P_{s}) &= Y_{t} X_{t} - Y_{0}
    X_{0} + \int_{0}^{t} Y_{s} c_{s} ds \\
    &\geq -Y_{t} L_{t} - Y_{0} X_{0}
  \end{align}  if $(H, c)$ is $L$-admissible. and from the lemma.

  Also, since $YP$ is a local martingale then $\int H \cdot d(YP)$ is
  a local martingale (by construction of the \ito integral), so there
  exists a sequence of stopping times $\tau_{n} \uparrow \infty$ such
  that $(\int H \cdot d(YP))^{\tau_{n}}$ is a true martingale.

  Then
  \begin{align}
    \label{eq:64}
    \E{\int_{0}^{T} H_{s} \cdot d(Y_{s} P_{s}) + Y_{T}L_{T}} &=
    \E{\lim_{n \rightarrow \infty} \int_{0}^{T \wedge \tau_{n}} H_{s}
      \cdot d(Y_{s} P_{s}) + L_{T \wedge \tau_{n}} Y_{T \wedge
        \tau_{n}} } \\
    \label{eq:65}
    &\leq \liminf_{n \rightarrow \infty} \E{\int_{0}^{T \wedge
        \tau_{n}} H d(YP) + L_{T \wedge \tau_{n}} Y_{T \wedge
        \tau_{n}}} \\
    \label{eq:66}
    &= \liminf_{n \rightarrow \infty} \E{Y_{T \wedge \tau_{n}} L_{T
        \wedge \tau_{n}}} \\
    \label{eq:67}
     &= \E{Y_{T} L_{T}}
  \end{align} by Fatau's lemma \eqref{eq:65}, using that $(\int_{0}^{t} H \cdot d(YP))^{\tau_{n}}$ is a
  martingale starting at zero \eqref{eq:66} and the assumption of
  uniform integrability \eqref{eq:67}.

  So suppose $X_{0} = 0 = X_{T}$ almost surely. Then
  \begin{align}
    \label{eq:68}
    \E{\int_{0}^{T} Y_{s} c_{s} ds} = \E{\int_{0}^{T} H_{s} \cdot
      d(Y_{s} P_{s})} \leq 0 \Rightarrow c_{t}(\omega) = 0 a.e.
  \end{align} which implies no arbitrage.
\end{proof}


Suppose $P = (N, S)$ where $N_{t} > 0$ for all $t \geq 0$ almost
surely - e.g. the price of a numeraire.

\begin{defn}
  \label{defn:continuous_time:27}
  A pure investment strategy $H$ is an arbitrage relative to the
  numeraire if and only if
  \begin{enumerate}
  \item There exists a non-random $T > 0$ such that
    \begin{equation}
      \label{eq:69}
      \frac{X_{T}}{N_{0}} \geq \frac{N_{T}}{N_{0}} a.s.
    \end{equation} and
    \begin{equation}
      \label{eq:70}
      \Prob{\frac{X_{T}}{N_{0}} > \frac{N_{T}}{N_{0}}} > 0
    \end{equation}
  \end{enumerate}
\end{defn}

\begin{remark}
  There exists a model $P$, credit limit $L$ such that there is no
  absolute arbitrage but there is a relative arbitrage. \todo{To show}
\end{remark}

\begin{defn}
  \label{defn:continuous_time:28}
  An equivalent (local) martingale measure is a measure $\Q \sim
  \Prob$ such that $\frac{S}{N}$ is a $\Q$-local martingale.
\end{defn}

\begin{thm}[FTAP1 for market with a numeraire]
  \label{defn:continuous_time:29}
  Suppose $\Q$ is an EMM and $\frac{L}{N}$ is locally class $D$ (with
  respect to $\Q$), then there are no $L$-admissible relative
  arbitrages.
\end{thm}

\begin{lem}
  If $X_{t}= \phi_{t} N_{t} + \pi_{t} \cdot S_{t}$ (i.e $(\psi, \pi)$
  is a self-financing pure investment stratey), then
  \begin{equation}
    \label{eq:71}
    d\frac{X_{t}}{N_{t}} = \pi_{t} d \frac{S_{t}}{N_{t}}.
  \end{equation}

\end{lem}

\begin{proof}
  Ito's lemma
\end{proof}

\begin{proof}[Proof of theorem]
  If $\Q$ is an EMM, $X$ is a $\Q$-local martinagle, since it is the
  stochastic integral with respect to the $Q$-local martingale
  $\frac{S}{N}$.  As $\frac{X_{t} + L_{t}}{N_{t}} \geq 0$, we can
  apply Fatao's lemma as before, obtaining
  \begin{align}
    \label{eq:72}
    \E{\frac{X_{T}}{N_{T}}}{Q} \leq \frac{X_{0}}{N_{0}}.
  \end{align}  Thus, if
  \begin{align}
    \label{eq:73}
    \frac{X_{T}}{N_{T}} \geq \frac{X_{0}}{N_{0}}
  \end{align} $\Prob$ a.s.  then
  \begin{align}
    \label{eq:74}
    \frac{X_{T}}{N_{T}} \geq \frac{X_{0}}{N_{0}}
  \end{align} $\Q$ a.s by equivalence of $\Prob$ and $\Q$.

  Then $\frac{X_{T}}{N_{T}} = \frac{X_{0}}{N_{0}}$ $\Q$ a.s. by the
  pigeon hole, then $\frac{X_{T}}{N_{T}} = \frac{X_{0}}{N_{0}}$
  $\Prob$ a.s, since $\Prob \sim \Q$.
\end{proof}

\todo{Fill in rest of lecture content}

In the framework $P = (B, S)$, $dB_{t} = B_{t} r_{t} dt$, $dS^{i}_{t} =
S_{i}(\mu^{i}_{t} dt + \sum_{j=1}^{m} \sigma^{ij}_{t} dW^{j}_{t})$. 

\begin{thm}
  \label{defn:continuous_time:30}
  Let $\lambda_{t}$ be predictable and $\int_{0}^{t} \| \lambda_{s}
  \|^{2} ds < 0$ a.s. $\forall t \geq 0$ and satisfying $\sigma_{t}
  \lambda_{t} = \mu_{t} = r_{t}$.  Then $dY_{t} = -Y_{t}(r_{t} dt +
  \lambda_{t} dW_{t})$ is a state price density and if $W$ generates
  the filtration then all state price densities are of this form.
  $\lambda$ is called a market price of risk.
\end{thm}

\begin{proof}
  From \ito's formula,
  \begin{align}
    \label{eq:75}
    d(Y_{t} B_{t}) = -Y_{t} B_{t} \lambda_{t} \cdot dW_{t}
  \end{align} is a local martingale,

  \begin{align}
    \label{eq:155}
    d(Y_{t} S_{t}^{i}) &= Y_{t}S_{t}^{i}(\mu^{i}_{t} + \sum \sigma^{ij}
    dW^{j}) + YS^{i}(-r dt - \sum \lambda^{j} dW^{j}) - Y S^{i} \sum
    \sigma^{ij} \lambda^{j} dt \\
    d(YS^{i}) &= YS^{i}((\sigma^{ij} - \lambda) dW + (\mu^{i} - r -
    (\sigma \lambda)^{i} dt))
  \end{align}

  Now, if the filtration is generated by $W$, then all positive local
  martingales $M$ are of the form (by the martingale representation
  theorem) $dM = -M \lambda \cdot dW$ for some predictable process
  $\lambda$.  So if $Y$ is a state price density then $Y$ is of the
  form $Y = \frac{M}{S}$ so $dY = - Y(rdt = \lambda dW)$.  If $YS^{i}$
  is a local martingale for all $i$ then $\sigma \lambda = u - r1$ in
  order for the $dt$ to cancel in \ito's formula.
\end{proof}

If $Y$ is a state price density such that $YB$ is a true martingale,
we can define an equivalent measure $\Q$ by $\frac{d\Q}{d\Prob} =
\frac{Y_{T} B_{T}}{Y_{0} B_{0}}$ for some fixed $T > 0$.  This $\Q$ is
an equivalent martingale measure.

\begin{thm}
  \label{defn:continuous_time:31}
  Suppose $dM_{t} = -M_{t} \lambda_{t} \cdot dW_{t}$ is a true
  martingale where $\lambda$ solves $\sigma \lambda = \mu - r1$.  Fix
  $T > 0$ and let $\frac{d\Q}{d\Prob} = \frac{M_{T}}{M_{0}}$.  Then
  $\Q$ is an EMM and $dS_{t}^{i} = S_{t}^{i}(r_{t} dt + \sigma^{ij}
  d\hat W_{t})$ for a $\Q$-Brownian motion $\hat W$.
\end{thm}

\begin{proof}
  By Girsanov's theorem, $\hat W_{t} = W_{t} + \int_{0}^{t}
  \lambda_{s} ds$ is a $\Q$-Brownian motion.  Now, by \ito,
  \begin{align}
    \label{eq:77}
    d(\frac{S_{i}}{B}) = \frac{S_{i}}{B}((\mu^{i} - r) dt +
    \sigma^{ij} dW) \\
    = \frac{S^{i}}{B} \sigma^{ij} (\lambda_{t} dt + dW_{t}) \\
    = \frac{S^{i}}{B} \sigma^{ij} d \hat W.
  \end{align}
\end{proof}

\begin{thm}
  \label{defn:continuous_time:32}
  Suppose that the filtration is generated by $W$.  Suppose $n = d$
  and that the $d \times d$ matrix $\sigma^{ij}(\omega)$ is invertible
  for all $t, \omega$.  Let $\lambda_{t} = \sigma_{t}^{ij}(\mu_{t} -
  r_{t} 1)$ and $dY_{t} = -Y_{t}(r_{t} dt + \lambda_{t} dW_{t})$ is
  the unique state price density.  Let $\xi_{T}$ be a
  $\mathcal{F}_{t}$-measurable non-negative random variable such that
  $\xi_{T} Y_{T}$ is integrable.  Then there exists a $0$-admissible
  trading strategy $H$ such that $X_{T}^{H} = \xi_{T}$ and $X_{0}^{H}
  = \frac{\E{Y_{T} \xi_{T}}}{Y_{0}}$.

  Furthermore, if $LY$ is locally of class $D$ and $\tilde H$ is an
  $L$-admissible strategy such that $X_{T}(\tilde H) = \xi_{T}$, then
  $X_{0}(\tilde H) \geq X_{0}(H)$.  That is, $\frac{\E{Y_{T}
      \xi_{T}}}{\xi_{0}}$ is the minimal replication cost of the
  European claim with payout $\xi_{T}$.
\end{thm}

\begin{proof}
  Let $M_{t} = \E{Y_{T} \xi_{T} | \mathcal{F}_{t}}$.  This is a
  martingale.  We show that there eixsts $H$ such that $X_{t}^{H} =
  \frac{M_{t}}{Y_{t}}$for all $0 \leq t \leq T$.  By the martingale
  representation theorem, there exists a $d$-dimensional predictable
  process $\alpha$ such that
  \begin{align}
    \label{eq:78}
    dM_{t} = \alpha_{t} dW_{t}
  \end{align}  By \ito's formula,
  \begin{align}
    \label{eq:79}
    d \frac{M_{t}}{Y_{t}} = \frac{M_{t}}{Y_{t}} r_{t} dt + (
    \frac{M_{t} \lambda_{t} + \sigma_{t}}{Y_{t}}) (dW_{t} +
    \lambda_{t} dt).
  \end{align}

  Let $\pi_{t} = \diag (S_{t})^{-1} (\sigma_{t}^{T})^{-1} (\frac{M_{t}
    \lambda_{t} + \sigma_{t}}{Y_{t}})$ and
  \begin{align}
    \label{eq:81}
    \phi_{t} = \frac{\frac{M_{t}}{Y_{t}} - \pi_{t} S_{t}}{B_{t}}.
  \end{align}  Note that $\phi_{t} B_{t} + \pi_{t} S_{t} =
  \frac{M_{t}}{Y_{t}}$, and
  \begin{align}
    \label{eq:82}
    \pi_{t} dB_{t} + \pi_{t} dS_{t} = \frac{M_{t}}{Y_{t}} r dt +
    \frac{M_{t} \lambda_{t} + \alpha}{Y_{t}} (dW  +\lambda dt) = d(\frac{M}{Y})
  \end{align} and so $H = (\phi, \pi)$ is a self-financing strategy.
  It is 0-admissible since $\frac{M_{t}}{Y_{t}} > 0$.
\end{proof}

\begin{thm}
  \label{defn:continuous_time:33}
  If $\tilde H$ is $L$-admissible and $LY$ is in class $D$ and
  $X_{T}(\tilde H) = \xi_{T}$ then
  \begin{align}
    \label{eq:80}
    X_{0}(\tilde H) \geq \frac{\E{Y_{T} \xi_{T}}}{Y_{0}}= X_{0}(H)
  \end{align}
\end{thm}

\begin{proof}
  Consider
  \begin{equation}
    \label{eq:83}
    -Y_{t}(\tilde X_{t} + L_{t}) \geq 0
  \end{equation} and $Y_{t} \tilde X_{t}$ is a local martingale.
  \begin{align}
    \label{eq:84}
    \E{Y_{T \wedge \tau_{n}} L_{T \wedge L_{n}}} \rightarrow \E{Y_{T}
      L_{T}}
  \end{align} by uniform integrability assumption. Therefore $Y \tilde
  X$ is a supermartinagle by Fatau's lemma, and thus
  \begin{align}
    \label{eq:85}
    \E{Y_{T}\xi_{T}} = \E{Y_{T} \tilde X_{T}} \leq Y_{0} \tilde X_{0}
  \end{align}
\end{proof}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "master"
%%% End:
