\begin{defn}
  \label{defn:new_convex:2}
  $f$ is lower semicountinuous if
  \begin{equation}
    \label{eq:6}
    f(x') \leq \liminf_{x \rightarrow x'} f(x) = \min \{\alpha \in \bar
    \R | (x^{k}) \rightarrow x_{1}', f(x^{k}) \rightarrow \alpha \}
  \end{equation}

\end{defn}

\begin{thm}
  \label{defn:new_convex:1}
  $f: R^{n} \rightarrow \bar \R$.  Then the following are equivalent
  \begin{enumerate}
  \item $f$ is lower semi continuous,
  \item $epi(f)$ is closed $C$ in $\R^{n} \times \R$
  \item $loevi(f) = \{ x \in \R^{n} | f(x) \leq \alpha \}$ are closed
    for all $\alpha \in \bar \R$
  \end{enumerate}
\end{thm}

\begin{proof}
  1 => 2 take $(x^{k}, \alpha^{k}) \in epi(f) \rightarrow (x, \alpha)
  \in \R^{n} \times \R$.  Then By 2.8 $f(x) \leq \liminf_{k
    \rightarrow \infty} f(x^{k}) \leq liminf_{k \rightarrow \infty}
  x^{k} = \alpha$ as $(x, \alpha) \in epi(f)$.

  2 => 3 $epi(f)$ closed implies $epi(f) \cap (\R^{n} \times \{ a'
  \})$ closed for all $\alpha' \in \R$, if and only if
  \begin{equation}
    \label{eq:7}
    \{ x \in \R^{n} | f(x) \leq \alpha' \}
  \end{equation} is closed for all $\alpha' \in \R$
  If $\alpha' = \infty$, $lev_{\leq +\infty}f = \R^{n}$.
  If $\alpha' = -\infty$, $lev_{\leq -\infty} f = \cap_{k \in \N}
    lev_{\leq -k} f$ which is the intersection of closed sets.
  
  3 => 2 for $x^{j} \rightarrow x'$, take the subsequence $x^{k}
  \rightarrow x'$, with
  \begin{equation}
    \label{eq:8}
    f(x^{k}) \rightarrow \liminf_{x \rightarrow x'} f(x) = c
  \end{equation}

  if $c \in \R$, for all $K(\epsilon) = K'$ such that $f(x^{k}) \leq C
  + \epsilon$ for all $k > K'$. Then

  $\Rightarrow x^{k} \in lev_{\leq C + \epsilon} f \Rightarrow x' \in
  lev_{\leq C + \epsilon}$.

  Equivalently, $x' \in lev_{\leq C} f \Rightarrow f(x') \leq C =
  \liminf{x \rightarrow x'} f(x)$.

  If $c = +\infty$, done - $f(x') \leq +\infty = \liminf$.
  If $c = \infty$, same argument with $k \in \N$, $f(x^{k}) \leq -k, ..$.
\end{proof}


$f(x) = x$ is lower semi-continuous, but does not have a minimizer.
\begin{defn}
  \label{defn:new_convex:3}
  $f: R^{n} \rightarrow \bar \R$ is \textbf{level-bounded} if and only
  if $lev_{\leq \alpha} f$ is bounded for all $\alpha \in \R$.

  This is also known as \textbf{coercivity}.
\end{defn}

\begin{exmp}
  \label{defn:new_convex:4}
  $f(x) = x^{2}$ is level-bounded.

  $f(x) = \frac{1}{|x|}$ is not level bounded.
\end{exmp}

\begin{thm}
  \label{defn:new_convex:5}
  $f: \R^{n} \rightarrow \bar \R$ is level bounded if and only if
  $f(x^{k}) \rightarrow \infty$ if $\| x^{k} \| \rightarrow \infty$.
\end{thm}

\begin{proof}
  Exercise.
\end{proof}


\begin{thm}
  \label{defn:new_convex:6}
  $f: \R^{n} \rightarrow \bar \R$ is lower semincontiuous,
  level-bounded, and proper. Then $\inf_{x} f(x) \in (-\infty,
  +\infty)$ and $argmin f = \{ x | f(x) \leq inf f(x) \}$ is nonempty and compact.
\end{thm}

\begin{proof}
  \begin{align}
    \label{eq:9}
    arg min f &= \{ x | f(x) \leq \inf_{x} f(x) \} \\
    &= \{ x | f(x) \leq \inf f(x)  + \frac{1}{k}, \forall k \in \N \}
  \\
  &= \cap_{k \in \N} lev_{\inf f + \frac{1}{k}} f
\end{align}

 If $\inf f$ is $-\infty$, just replace $\inf f + \frac{1}{k}$ by
 $\alpha$ with $\alpha > \infty$, and set $\alpha_{k} = -k, k \in \N$.
 
 These are bounded (by level-boundedness), closed (by f
 being lower semicontinous), and non-empty (since $\frac{1}{k} \geq
 0$.).  Then these limit sets are compact, we can just take the limit
 of the left boundaries of the level sets, and construct the
 convergent subsequence that is contained in every level set.

 We need to show that $\inf f \neq -\infty$.  If $\inf f = -\infty$,
 then there exists $x \in argmin f$ with $f(x) = -\infty$.  Since $f$
 is proper, this cannot exist. Thus, $\inf f \neq = -infty$.
\end{proof}

\begin{remark}
  For Theorem \ref{defn:new_convex:6}, it suffices to have $lev_{\leq
    \alpha} f$ bounded and non-empty for at least one $\alpha \in \R$. 
\end{remark}

\begin{proposition}
  \begin{enumerate}
  \item If $f,g $ is lower semicontinous, then $f + g$ is lower semicontinous
  \item IF $f$ is lower semicontinous, $\lambda \geq 0$, then
    $\lambda f$ is lower semicontinuous
  \item $f: \R^{n} \rightarrow \bar \R$ is lower semicontinuous, $g:
    \R^{m} \rightarrow \R^{n}$ is continous, then $f \circ g$ is lower semicontinous.
  \end{enumerate}
\end{proposition}



%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "master"
%%% End: 
